Evaluation of a Cache-Oblivious Data Structure [draft-1]

### 2008-05-31 paper/paper.tex
Updates to the paper draft.

When it comes to memory usage the B-tree and hash table have similar requirements;
the B-tree has slightly larger overhead per value stored, but initially the hash
table uses more memory because of the space allocated for the index.

@@

Finally, it should be noted that in our experiments only part of the memory
hierarchy was used (up to the use of main memory). This still involves several
levels of processor cache, but it is not a very deep hierarchy. With a deeper
memory hierarchy, with greater differences in access times between the levels,
the cache-efficient data structures (the B-tree and the Bender set) should
perform relatively better.

@@

>>>>

With respect to memory usage the B-tree and hash table have similar requirements;
the B-tree has slightly larger overhead per value stored, but initially the hash
table uses more memory because of the space allocated for the index.

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{ l r r }
\hline
\textbf{Purpose} & \textbf{Lines} & \textbf{Percentage} \\
\hline
Common code       & 433 & 18.21\% \\
Hash table        & 146 &  6.14\% \\
B-tree            & 299 & 12.57\% \\
Bender set        & 623 & 26.20\% \\
Queue             & 184 &  7.74\% \\
Search Framework  & 693 & 29.14\% \\
\hline
\end{tabular}
\caption{Source lines of code of the test application}
\label{tab-loc}
\end{center}
\end{table}

Table \ref {tab-loc} gives an overview of the source lines of code used for
the test application, after removing comments and blank lines (which together
amount for about 25\% of the code). Common code includes allocation functions,
interface descriptions and comparison and hashing functions. The framework
includes not just the search algorithm, but also the functionality to report
various metrics while running a test case.

Although not a perfect metric of implementation complexity, the lines of code
required for the various components of the test application do give some
insight in the relative complexity of the data structures. It is apparent why
hash tables are popular: they perform well while they are relatively easy to
implement.

\section{Future work}
[[Measure different data structures]]

[[Measure performance in low memory environments (using disk)]]

[[Mesure cache performance specifically]]

It should be noted that in our experiments only part of the memory
hierarchy was used (up to the use of main memory). This still involves several
levels of processor cache, but it is not a very deep hierarchy. With a deeper
memory hierarchy, with greater differences in access times between the levels,
the cache-efficient data structures (the B-tree and the Bender set) should
perform relatively better.
